---
title: "p8105_hw2_yl4925"
author: "Yiming Li"
date: "10/3/2021"
output: github_document
---

## Problem 1
```{r, include = FALSE}
library(tidyverse)
```

Load trash wheel as required
```{r}
trash_wheel_df = readxl::read_excel(
  "./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "Mr. Trash Wheel",
  range = "A2:N535") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  filter(dumpster != "Grand Total") %>% 
  mutate(sports_balls = round(sports_balls))
trashrow = nrow(trash_wheel_df)
trashcol = ncol(trash_wheel_df)
```


Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year. Next, combine precipitation dataset and convert month to a character variable
```{r, message = FALSE}
pre2019 = readxl::read_excel(
  "./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "2019 Precipitation",
  range = "A2:B14"
) %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2019) %>% 
  relocate(year)

pre2018 = readxl::read_excel(
  "./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "2018 Precipitation",
  range = "A2:B14"
) %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
pre2019row = nrow(pre2019)
pre2019col = ncol(pre2019)
pre2018row = nrow(pre2018)
pre2018col = ncol(pre2018)
```

```{r, message = FALSE}
##month.name is a vector, not function
bind_rows(pre2018, pre2019) %>% 
  mutate(month = month.name[month])
```

```{r}
sum_2018 = sum(pull(pre2018, total))
ball_2019 = filter(trash_wheel_df, year == 2019)
median_2019 = median(pull(ball_2019, sports_balls))
```


```{r, include = FALSE}
wtmin = min(pull(trash_wheel_df, weight_tons))
wtmax = max(pull(trash_wheel_df, weight_tons))
wtmean = mean(pull(trash_wheel_df, weight_tons))
vcymin = min(pull(trash_wheel_df, var = 6))
vcymax = max(pull(trash_wheel_df, var = 6))
vcymean = mean(pull(trash_wheel_df, var = 6))
pbmin = min(pull(trash_wheel_df, var = 7))
pbmax = max(pull(trash_wheel_df, var = 7))
pbmean = mean(pull(trash_wheel_df, var = 7))
pmin = min(pull(trash_wheel_df, var = 8))
pmax = max(pull(trash_wheel_df, var = 8))
pmean = mean(pull(trash_wheel_df, var = 8))
```

Trash Wheel size is (`r trashrow`, `r trashcol`). When we load Trash Wheel data, we find that it includes dumpster labels, date variable(month, year, date) and totally 10 different categories of trash(Weight_tons, volume_cubic_yards, plastic_bottles, polystyrene, cigarette_butts, glass_bottles, grocery_bags, chip_bags, sports_bags, homes_powered). Weight_tons ranges from `r wtmin` to `r wtmax` with mean `r wtmean`, volume_cubic_yards ranges from `r vcymin` to `r vcymax` with mean `r vcymean`, plastic_bottles ranges from `r pbmin` to `r pbmax` with mean `r pbmean`, polystyrene ranges from `r pmin` to `r pmax` with mean `r pmean`. (other types of trash are omitted)

The size for 2018 precipitation and 2019 precipitation both are (`r pre2018row`, `r pre2018col`).
The total precipitation in 2018 is `r sum_2018`, the median number of sports balls in a dumpster in 2019 is `r median_2019`.

## Problem 2

Load pols_month as required, add president variable to show party of president. 
```{r, message = FALSE}
pols_month = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(month = month.name[as.integer(month)],
         president = if_else(prez_gop == 1, "republic", "democratic")) %>% 
  select(-day, -prez_gop, -prez_dem)
```

Load snp data and rearrange data through 2-digit year
```{r, message = FALSE}
snp_raw = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  relocate(year, month) %>% 
  mutate(month = as.integer(month)) %>% 
  arrange(year, month) %>% 
  mutate(month = month.name[month]) %>%
  select(-"day")

snp_after_2000 = filter(snp_raw, year <= 20) %>% 
  mutate(year = as.character(as.numeric(year) + 2000))
snp_before_2000 = filter(snp_raw, year >= 20) %>% 
  mutate(year = as.character(as.numeric(year) + 1900))
snp = bind_rows(snp_before_2000, snp_after_2000) 
```

Load unemployment data and switch it to longer version.
```{r, message = FALSE}
unemployment = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec, 
    names_to = "month",
    values_to = "unemployment"
  ) %>% 
  mutate(month = month.name[match(month,month.abb)]) %>% 
  rename(year = Year) %>% 
  mutate(year = as.character(year))
```

Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r, message = FALSE}
pols_snp = left_join(pols_month, snp) 
pols_snp_unemployment = left_join(pols_snp, unemployment)
```

pols_month is dataframe for number of national politicians, size is (`r dim(pols_month)`). It contains number of democratic governors, senators, representatives and presidents, number of republic governors, senators, representatives and presidents. And it contains date from Jan 1947 - June 2015.(variables contain year, month, gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, rep_dem)


snp is dataframe for Standard & Poor’s stock market index, size is (`r dim(snp)`). It contains the close index for S&P and date from Jan 1950 to July 2015.(variables contain year, month, close)

unemployment is dataframe for unemployment, size is (`r dim(unemployment)`). It contains the percentage of unemployment and date from Jan 1948 to June 2015.(variables contain year, month, unemployment)

pols_snp_unemployment join these three datasets, so it includes all variables above. Its dimension is (`r dim(pols_snp_unemployment)`). Its date is from Jan 1947 to June 2015(same to pols_month)

## Problem 3
Load and tidy the data. (notice that ethnicity and name in 2012 have same problems )
```{r, message = FALSE}
## 2012是bug的一年，名字和缩写都不对
pop_baby_name = read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                            "BLACK NON HISP" = "BLACK NON HISPANIC",
                            "WHITE NON HISP" = "WHITE NON HISPANIC"),
         childs_first_name = str_to_sentence(childs_first_name),
         gender = str_to_sentence(gender),
         ethnicity = str_to_sentence(ethnicity))
```

delete duplicated row based on year, gender, ethnicity and first name.
```{r}
duplictaed_ones = pop_baby_name[duplicated(pop_baby_name[,1:4]),]
distinct_pop_name = pop_baby_name[!duplicated(pop_baby_name[,1:4]),]
```

Female Olivia rank in different ethnicity among years.
```{r}
Olivia_female_rank = filter(distinct_pop_name, gender == "Female" & childs_first_name == "Olivia") %>% 
  select(-gender, -childs_first_name, -count) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank
  )
```

Most popular male name in different ethnicity among years.
```{r}
most_popular_male_name = filter(distinct_pop_name, gender == "Male" & rank == 1) %>% 
  select(-gender, -rank, -count) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = childs_first_name
  )
```

produce a scatter plot showing the number of children with a name against the rank in popularity of that name for male white non hispanic baby in 2016.
```{r}
male_nonhisp_2016 = filter(distinct_pop_name, gender == "Male", 
                           ethnicity == "White non hispanic", 
                           year_of_birth == 2016)
##library(ggrepel)
ggplot(male_nonhisp_2016, aes(x = rank, y = count)) + 
  labs(
    title = "2016 white non-hispanic male children's name and rank", 
    x = "rank",
    y = "number of children",
  ) +
  geom_point() 
##+ geom_text_repel(aes(x = rank, y = count, label = childs_first_name))
```

